Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-12-20 ms.tex
Finishing Shane's edits

We now compare each of the previously-discussed solutions for surface flux to the Monte Carlo results. The spectrum $P(x)$ is defined as the specific luminosity at the surface divided by the source luminosity, or
\be \label{eq:prob_spectrum}
P(x) = \frac{16\pi^2R^2H(R, x)\Delta}{L}.
\ee
This is normalized as $\int P(x)dx = 1$. Since $H(R, x)$ is per $d\nu$, a factor of $\Delta$ gives the expression the correct units.

@@

In Figure \ref{fig:escape_time} escape time distributions calculated from Equation (\ref{eq:waittime}) are shown alongside Monte Carlo and the fitting function Equation (\ref{eq:fitting_function}) for $\tau_0=10^6, 10^7$ with $x_s=0$. It is seen that the fitting function matches well with the late-time behavior of the eigenfunction solution, indicating that the tail of the escape time distribution can be modeled effectively by simple exponential decay. The remaining disagreement between the tail of the distribution and the Monte Carlo data is due to line core scattering which is not modeled by the eigenfunction solution, but improves for larger optical depth as seen in the figure. Many scatterings in the Doppler core affects the tail of the escape-time distribution, since photons with frequencies near line center will take longer to escape. Thus, the rate constant for the exponential falloff is overestimated slightly in the eigenfunction solution as compared with the Monte Carlo. The error in this rate constant is a function of $\tau_0$ since the effect from the Doppler core is greatest when it extends into the peak of the spectrum.

@@

This probability distribution can be used to define the moments of the frequency shift
\be
\langle \delta \nu^n \rangle & = & \frac{ \int d\nu^\prime R (\nu^\prime-\nu)^n}{\int d\nu^\prime R}
= \frac{1}{\phi(\nu)}
\int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\,
\left( \frac{\nu_0 \vec{v} \cdot (\vec{n}^\prime-\vec{n}) }{c} \right)^n
\left( \frac{\Gamma/4\pi^2}{ \left(\nu - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}/c \right)^2 + (\Gamma/4\pi)^2 } \right),
\ee
which are functions of $\nu$, $\vec{n}$ and $\vec{n}^\prime$. These integrals can be evaluated in terms of the dimensionless moments of the parallel velocity distribution, defined as
\be
\langle u_\parallel^n \rangle(x,a) & = & \frac{a/\pi }{\mathcal{H}(x,a)} \int 
\frac{du_\parallel u_\parallel^n e^{-u_\parallel^2}  }{(x-u_\parallel)^2 + a^2}.
\ee
The end results for the first and second moments are
\be
\langle \delta \nu \rangle & = & -\langle u_\parallel \rangle \left( 1 - \vec{n} \cdot \vec{n}^\prime \right) \Delta 
\\
\langle \delta \nu^2 \rangle & = & 
\left[ \langle u_\parallel^2 \rangle
\left( 1 - \vec{n} \cdot \vec{n}^\prime \right)^2
+ \frac{1}{2} \left( 1 - \left( \vec{n} \cdot \vec{n}^\prime\right)^2 \right) \right] \Delta^2 .
\ee

>>>>

We now compare each of the previously-discussed solutions for surface flux to the Monte Carlo results. The spectrum $P(x)$ is defined as the specific luminosity at the surface divided by the source luminosity, or
\be \label{eq:prob_spectrum}
P(x) = \frac{16\pi^2R^2H(R, x)\Delta}{L}.
\ee
This is normalized so that $\int P(x)dx = 1$. Since $H(R, x)$ is per $d\nu$, a factor of $\Delta$ gives the expression the correct units.

@@

In Figure \ref{fig:escape_time} escape time distributions calculated from Equation (\ref{eq:waittime}) are shown alongside Monte Carlo and the fitting function Equation (\ref{eq:fitting_function}) for $\tau_0=10^6, 10^7$ with $x_s=0$. It is seen that the fitting function matches well with the late-time behavior of the eigenfunction solution, indicating that the tail of the escape time distribution can be modeled effectively by simple exponential decay. The remaining disagreement between the tail of the distribution and the Monte Carlo data is due to line core scattering which is not modeled by the eigenfunction solution, but improves for larger optical depth as seen in the figure. A large number of scatterings in the Doppler core affects the tail of the escape-time distribution, since photons with frequencies near line center will take longer to escape. Thus, the rate constant for the exponential falloff is overestimated slightly in the eigenfunction solution as compared with the Monte Carlo. The error in this rate constant is a function of $\tau_0$ since the effect from the Doppler core is greatest when it extends into the peak of the spectrum.

@@

This probability distribution can be used to define the moments of the frequency shift
\be
\langle \delta \nu^n \rangle & = & \frac{ \int d\nu^\prime (\nu^\prime-\nu)^n R}{\int d\nu^\prime R}
= \frac{1}{\phi(\nu)}
\int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\,
\left( \frac{\nu_0 \vec{v} \cdot (\vec{n}^\prime-\vec{n}) }{c} \right)^n
\left( \frac{\Gamma/4\pi^2}{ \left(\nu - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}/c \right)^2 + (\Gamma/4\pi)^2 } \right),
\ee
which are functions of $\nu$, $\vec{n}$ and $\vec{n}^\prime$. These integrals can be evaluated in terms of the dimensionless moments of the parallel velocity distribution, defined as
\be
\langle u_\parallel^n \rangle(x,a) & = & \frac{a/\pi }{\mathcal{H}(x,a)} \int 
\frac{du_\parallel u_\parallel^n e^{-u_\parallel^2}  }{(x-u_\parallel)^2 + a^2}.
\ee
The end results for the first and second moments are
\be
\langle \delta \nu \rangle & = & -\langle u_\parallel \rangle \left( 1 - \vec{n} \cdot \vec{n}^\prime \right) \Delta 
\\
\langle \delta \nu^2 \rangle & = & 
\left[ \langle u_\parallel^2 \rangle
\left( 1 - \vec{n} \cdot \vec{n}^\prime \right)^2
+ \frac{1}{2} \left( 1 - \left( \vec{n} \cdot \vec{n}^\prime\right)^2 \right) \right] \Delta^2 .
\ee
